---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---

## Loading and preprocessing the data
The First step is to load the data. We will create  a temporary file, download the zip file and extract it into mydata data frame object.
```{r Question1}
temp_data <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata/data/activity.zip",temp_data,method="curl")
mydata <- read.csv(unz(temp_data, "activity.csv"),sep=",",na.strings="NA")
mydata$sdate<-strptime(mydata$date,"%Y-%m-%d")
unlink(temp_data)
str(mydata)
```

## What is mean total number of steps taken per day?
```{r Question2}
library(doBy)
sumedData <- summaryBy(steps ~ date, data=mydata, FUN=c(sum),na.rm=T)
hist(sumedData$steps.sum,xlab="Total number of steps (blue-mean; red-median)",ylab="Frequency",main="Histogram of Steps")
meanSteps <- round(mean(sumedData$steps.sum,na.rm=T), 2)
medianSteps <- round(median(sumedData$steps.sum,na.rm=T), 2)
abline(v=meanSteps, lwd = 3, col = 'blue')
abline(v=medianSteps, lwd = 3, col = 'red')
```

The daily Average steps is `r meanSteps` 
the daily median Steps is `r medianSteps`


## What is the average daily activity pattern?
```{r Question3}
IntervalSumedData <- summaryBy(steps ~ interval, data=mydata, FUN=c(mean),na.rm=T)
plot(IntervalSumedData$interval,IntervalSumedData$steps.mean,type="l",xlab="5 Minute Time Interval",ylab="Average Number of Steps",main="Average Steps by Time Interval")
maxInterval<-IntervalSumedData[which.max(IntervalSumedData$steps.mean),]
```

interval number `r which.max(IntervalSumedData$steps.mean)` have the maximun average steps. 
this the `r maxInterval$interval` minute with average of `r round(maxInterval$steps.mean,2)` steps

## Imputing missing values

Total number of missing values is

```{r Question4a}
sum(is.na(mydata$steps))
```

We replace the NA with the 5-minute interval average

```{r Question4b}
imputeData<-mydata
for (i in which(is.na(imputeData$steps))){
        imputeData$steps[i]<-round(IntervalSumedData$steps.mean[IntervalSumedData$interval==
                                                                        imputeData$interval[i]],0)
}
sum(is.na(imputeData$steps))
imputedSumedData <- summaryBy(steps ~ date, data=imputeData, FUN=c(sum),na.rm=T)
hist(imputedSumedData$steps.sum,xlab="Total number of steps (blue-mean; red-median)",ylab="Frequency",main="Histogram of Steps (Imputed Dataset)")
imputedMeanSteps <- round(mean(imputedSumedData$steps.sum,na.rm=T), 2)
imputedMedianSteps <- round(median(imputedSumedData$steps.sum,na.rm=T), 2)
abline(v=imputedMeanSteps, lwd = 3, col = 'blue')
abline(v=imputedMedianSteps, lwd = 3, col = 'red')
```

The daily Average steps in the imputed data set is `r imputedMeanSteps` compeared with `r meanSteps` in the non-imputed one.

the daily median Steps in the imputed data set is `r imputedMedianSteps` compeared with `r medianSteps` in the non-imputed one.

Answer To Question: The mean and the median are now the same and higher than in the original dataset.

## Are there differences in activity patterns between weekdays and weekends?
```{r Question5}
library(lattice)
Sys.setlocale("LC_TIME", "C")

isweekend = function(x){
        wday<-weekdays(strptime(x,"%Y-%m-%d"))
        if(wday %in% c('Saturday', 'Sunday')){
                return("weekend")
        }
        return("weekday")
}
imputeData$daytype = as.factor(apply(as.matrix(imputeData$date), 1, isweekend))
imputedIntervalSumedData <- summaryBy(steps ~ interval+daytype, data=imputeData, FUN=c(mean),na.rm=T)

xyplot(steps.mean~interval | daytype, data = imputedIntervalSumedData,
      type = 'l',
      xlab = 'Interval',
      ylab = 'Number of Steps',
      layout = c(1,2))
```
